### 2017-04-06 Ben_Manuscripts/XSEDE/bridges_2017/proposed_work.tex
Most of the content is there

subsection{How do water and ions move in nanostructured charged polymeric membranes?}
\subsubsection*{Background}

@@

The membrane system we propose to model in this project is composed of
lyotropic liquid crystals (LLCs) which self-assemble into hexagonally
packed, cylindrical pores with acidic groups facing toward the cylinder
center, complexed to counterions. Subsequent cross-linking to forms
a mechanically strong structure (Fig.~\ref{figure:membrane}).
~\cite{Zhou2003,Feng2014,Feng2016} The pores, at the level of resolution
obtained by SAXS and TEM, appear to be straight and uniform in size 
(Fig.~\ref{figure:membrane}).  This is very different from most 
commercially available membranes which have a pore size distribution
with tortuous pathways that reduce selectivity and flux respectively.
The ordered structure exhibited by LLC membranes makes them well suited 
for modeling using molecular dynamics simulations because the
system has a defined structure that can be studied in a reasonably sized 
unit cell.

@@

% BJC: The following is technically true although I've been creating
% specific configurations for testing purposes. 
In order to assemble the structures shown in Fig.~\ref{figure:membrane}, a
suitable monomer configuration needed to be chosen. Thermal annealing
of the monomer using GROMACS yielded many low energy configurations
close to what is expected.

Initial simulations have been performed which energy minimize and
allow the system to stabilize over the course of 20 ns in vacuum.  The
final structure and trajectory are analyzed quantitatively by
measuring pore size, distance between pores, and distribution of
sodium ions in the pore. These measurements are compared to values
measured experimentally with SAXS and TEM imaging in order to validate
the final structure. SAXS and TEM images were taken using dry membrane
which justifies our initial study of the system in vacuum.

%\begin{figure}[h]
%\begin{center}
%\begin{tabular}{c}
%\includegraphics[width=8cm]{modeled.png}\\
%\includegraphics[height=5cm]{membrane2.pdf}\\
%\end{tabular}
%\end{center}
%\caption{(above) SAXS and TEM experimental data, with hypothesized
%  membrane structure. (Below, left) Top view of atomistic simulation,
%  with hydrophilic head groups in orange, counterions in blue, and
%  hydrophobic chains in blue. (Below, right) Side view of atomistic
%  simulation showing counterions in dark blue and other atoms in light
%  blue. These simulations demonstrate the heterogeneity in the
%  modeled systems.~\label{figure:membrane}}
%\end{figure}

@@

We repeated scaling studies with the same system using GPU-enabled 
GROMACS 2016 which was compiled for optimal use with NVIDIA Tesla K80
accelerators. On Bridges we obtain 58 ns/day, 56.4 ns/day and 10.1 ns/day
for the 66K, 100K and 256K atom systems respectively. Although the system
scales nearly linearly for greater than 4 GPUs in all cases, additional
GPUs require more than 1 node which has signficantly extended queue wait
times in our experience. It is preferable to run jobs longer on one GPU
for a long time than to delay a job from starting but have it run for a
shorter time. Since trajectory files are written while simulating, it
is beneficial to check progress and cancel jobs which show no value.
We plan to work with PSC to compile a version of gromacs that is performs
optimally with the

In order to study the relative stability of the two metastable phases 
we have discovered, we need will need to perform a computationally 
intensive free energy calculation using the Multistate Bennett
Acceptance Ratio (MBAR) technique. MBAR estimates free energy 
difference with current lowest variance when compared to other estimators.
In order to make a reasonable estimate, we will need to conduct 
simulations all intermediate states which lead from one metastable
state to another. Each configuration in the pathway mapping the two
states must be sufficiently similar to adjacent configurations so that
we achieve enough phase space overlap for a more precise calculation. 
We estimate that we will need at least 50 intermediate states, each run
for 50 ns for a total of 2500 ns of simulation time. This will require 
(2500 ns / (99.2 ns/day))* 24 hrs/day * 224 cores = 135K SU using
MPI on the RM partition. This will require (2500 ns / (58 ns/day)) * 24
hours * 4 GPUs = 4.1K SU on the K80 GPU nodes.

The resolution of simulated X-ray diffraction patterns is dependent on the
size of the simulated unit cell. To create higher resolution patterns in 
the x, y or z directions, requires an increase in the respective dimension
of the unit cell by adding more atoms. The fundamental reason for this
limitation is the necessity of meeting the Bragg condition. When met,
constructive X-ray interference occurs, producing a signal on a nearby 
detector. The lattice planes in the crystal, defined by the reciprocal
space Miller indices h, k, and l, are separated by a distance, d. One can
calculate all possible d values given all unit cell parameters. It is
not trivial to see that an increase in box vector leads to a wider range
of accessible hkl values and increases the resolution. A simple way to 
estimate the simulated resolution in each direction is using the equations
qx = 2*pi/x, qy = 2*pi/y, qz = 2*pi/z. Our current resolution with a 66K
atom system is ~ 0.078 inverse angstroms. We propose a 4x increase in our
z dimension resolution which will help us to distinguish reflections 
deemed a consequence of benzene ring pi-stacking (reflection occurs at 1.53
inverse angstroms) from simple alkane chain packing (reflection occurs
between 1.4 and 1.57 inverse angstroms). We also hope to pick up finer 
details such as the sharp line that appears at ~.85 inverse angstroms 
experimentally but is absent in the simulated patterns. A system of this 
type is made of 265K atoms. Stacking equilibrated membrane layers 
directly on top of each will allow facilitate a fast equilibration of the
large system. Equilibration will be allowed to run for at least 50 ns 
followed by 50 ns of simulation needed to collect enough information to
simulate the XRD pattern. 100 ns of simulation time for a 265K atom system 
will require 19K SU on the RM nodes and 950 SU on K80 GPU nodes.     
%HIGH RESOLUTION XRD (to pick up additional features)

%CROSSLINKING?

Once equilibrated and cross-linked, we will examine the effect of
solvation of the membranes on the pores. We have learned that at least 
1000 ns of total simulation is required to observe full penetration of
water through pores resulting in a cost of 
%UPDATE
(2000 ns / (86 ns/day)) * 24
hours * 224 cores = 124K required SU, if we assume the initial
membrane simulation boxes are sufficiently large.
%UPDATE

We will carry out the same procedure with a set of new monomers. In addition to the simulations outlined above, we will need to equilibrate each system
using the developed procedures. Equilibration for 500 ns

>>>>

\subsection{How do water and ions move in nanostructured charged polymeric membranes?}
\subsubsection*{Background}

@@

The membrane system we propose to model in this project is composed of
lyotropic liquid crystals (LLCs) which self-assemble into hexagonally
packed, cylindrical pores with acidic groups facing toward the cylinder
center, complexed to counterions. Subsequent cross-linking of vinyl
groups on monomer tail ends forms a mechanically strong structure
(Fig.~\ref{figure:membrane}). ~\cite{Zhou2003,Feng2014,Feng2016}
The pores, at the level of resolution obtained by SAXS and TEM, appear
to be straight and uniform in size (Fig.~\ref{figure:membrane}). 
This is very different from most commercially available membranes 
which have a pore size distribution with tortuous pathways that reduce
selectivity and flux respectively. The ordered structure exhibited by
LLC membranes makes them well suited for modeling using molecular 
dynamics simulations because the system has a defined structure that
can be studied in detail with a reasonably sized unit cell.

@@

Equilibration simulations have been performed which energy minimize and
allow the system to stabilize over the course of 500 ns in vacuum.  The
final structure and trajectory are analyzed quantitatively by
measuring pore size, distance between pores, distribution of
sodium ions in the pore, simulating X-ray diffraction patterns and 
calculating ionic conductivity. These measurements are compared to values
measured experimentally with SAXS and TEM imaging in order to validate
the final structure. SAXS and TEM images were taken using dry membrane
which justifies our initial study of the system in vacuum.

\begin{figure}[h]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=8cm]{modeled.png}\\
\includegraphics[height=5cm]{membraneviews.png}\\
\end{tabular}
\end{center}
\caption{(above) SAXS and TEM experimental data, with hypothesized
  membrane structure. (Below, left) Top view of atomistic simulation,
  with hydrophilic head groups in orange, counterions in blue, and
  hydrophobic chains in blue. (Below, right) Side view of atomistic
  simulation showing counterions in dark blue and other atoms in light
  blue. These simulations demonstrate the heterogeneity in the
  modeled systems.~\label{figure:membrane}}
\end{figure}

@@

We repeated scaling studies with the same systems using GPU-enabled 
GROMACS 2016 with both the NVIDIA K80 and NVIDIA P100 GPUs available on
Bridges. Using 4 K80 GPUs we obtain 58 ns/day, 56.4 ns/day and 10.1 ns/day
for the 66K, 100K and 256K atom systems respectively. Although the system
scales nearly linearly for greater than 4 GPUs in all cases, additional
GPUs require more than 1 node which has signficantly extended queue wait
times in our experience. It is preferable to run jobs longer on one GPU
for a long time than to delay a job from starting but have it run for a
shorter time. Since trajectory files are written while simulating, it
is beneficial to check progress and cancel jobs which show no value.
We plan to work with PSC to compile a version of gromacs that is performs
optimally with the

In order to study the relative stability of the two metastable phases 
we have discovered, we will need to perform a computationally 
intensive free energy calculation using the Multistate Bennett
Acceptance Ratio (MBAR) technique. MBAR estimates free energy 
differences with the currently lowest variance when compared to other
estimators. In order to make a reasonable estimate, we will need to 
conduct simulations of all intermediate states which lead from one metastable
state to another. Each configuration in the pathway mapping the two
states must be sufficiently similar to adjacent configurations so that
we achieve enough phase space overlap for a more precise calculation. 
We estimate that we will need at least 50 intermediate states, each run
for 50 ns for a total of 2500 ns of simulation time. This will require 
(2500 ns / (99.2 ns/day))* 24 hrs/day * 224 cores = 135K SU using
MPI on the RM partition. This will require (2500 ns / (58 ns/day)) * 24
hours * 4 GPUs = 4.1K SU on the K80 GPU nodes.

The resolution of simulated X-ray diffraction patterns is dependent on the
size of the simulated unit cell. To create higher resolution patterns in 
the x, y or z directions, requires an increase in the respective dimension
of the unit cell by adding more atoms. The fundamental reason for this
limitation is the necessity of meeting the Bragg condition. When met,
constructive X-ray interference occurs resulting in a signal which gives
details about the position of atoms relative to each other. The lattice
planes in the crystal, defined by the reciprocal space Miller indices 
h, k, and l, are separated by a distance, d. One can calculate all possible
d values given all unit cell parameters. It is not trivial to see that an
increase in box vector leads to a wider range of accessible hkl values 
and increases the spatial resolution. A simple way to estimate the simulated
resolution in each direction is using the equations qx = 2*pi/x, qy = 2*pi/y,
qz = 2*pi/z. Our current resolution with a 66K atom system is ~ 0.078 inverse
angstroms (Fig.~\ref{fig:expxrdcomp}). We propose a 4x increase in our z dimension resolution which will
help us to distinguish reflections deemed a consequence of benzene ring 
pi-stacking (reflection occurs at 1.53 inverse angstroms) from simple
alkane chain packing (reflection occurs between 1.4 and 1.57 inverse
angstroms). We also hope to pick up finer details such as the sharp line
that appears at ~.85 inverse angstroms experimentally but is absent in 
the simulated patterns. A system of this size is made of 265K atoms. 
Stacking equilibrated membrane layers directly on top of each will 
facilitate a fast equilibration of the large system. Equilibration 
simulations will be run for at least 50 ns followed by another 50 ns of 
simulation needed to collect enough information to simulate the XRD 
pattern. 100 ns of simulation time for a 265K atom system will require 
19K SU on the RM nodes and 950 SU on K80 GPU nodes.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{Sim_exp_xrd_sidebyside.png}
	\caption{(left) Experimental Wide Angle X-ray Diffraction experiments contain
reflections at qz \approx 1.7 A $^{-1}$ indicating pi-stacking of benzene
rings in the monomer head group. A weak, correlated reflection is present at 
qz \approx 0.85 A $^{-1}$. (right) Simulated X-ray diffraction shows a pi-stacking reflection
at qz \approx 1.5 A $^{-1}$. The weak correlated reflection is not
present but may be visible with higher resolution simulations}
	\label{fig:expxrdcomp}

\end{figure}

Once equilibrated and cross-linked, we will examine the effect of membrane
solvation on the pore structure and corresponding transport properties.
We have learned that at least 1000 ns of total simulation is required
to fully equilibrate the system with water. (1000 ns / (86 ns/day)) * 24
hours * 224 cores = 62K required SU, if we assume the initial
membrane simulation boxes are sufficiently large.

We will carry out the same procedure with a new set of monomers similar
in structure to the current monomer. Simple modifications can be made
to the monomer structure as outlined in Figure ~\ref{fig:chemistries}. 
We will need to equilibrate each system using developed procedures which
requires equilibration simulations of at least 500 ns. Of the 
possibilities outlined in Figure ~\ref{fig:chemistries} we will try all
variations of tail spacer and ionic headgroup and at least 3 counterions
of different charge (+1, +2, +3). This results in a total of 24 systems 
that require 500 ns of equilbration time. These length simulations should
also give us the required information to calculate physical properties 
such as ionic conductivity, and to simulate X-ray diffraction patterns.
We will require xxx SU explore this space. To solvate all the test
systems, an additional xxx SU will be necessary.

We will conduct transport studies on all systems in order to relate pore
structure to macroscopic observables such as water flux and solute
rejection. We will measure diffusion of various solutes across the 
membrane. Understanding the mechanims of transport of each solute will
provide us with sufficient understanding to suggest new modifications 
to monomer structure that should increase laboratory performance. We
anticipate simulations of at least 50 ns will be necessary to get 
sufficient sampling for an informative mean square displacement (MSD) 
curve. The MSD curve will make it easier to identify potential 
transport mechanisms based on its shape.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{chemistries.png}  % who made this picture? It has a typo : "ionic headgoup"
	\caption{LLC chemistries we will examine in order to understand determinants of pore size, structure and solute transport}
	\label{fig:chemistries}
\end{figure}
